\speculake: Hiding Malware in Speculative Execution

### 2018-08-03 paper/implementation.tex
pass on implementation/payload reording

\section{Implementation and Evaluation}
We now discuss the details of our implementations and the information learned
about speculative execution gained by these experiments. We discuss how we gain
confidence in information passed back, and how a low number of iterations is
sufficient to gain a strong confidence in results.

For each of these payloads, we used our custom trigger program that performs 28
indirect jumps, followed by a final indirect jump to a function pointer that
pointed to a gadget at address 0x430000. In each of our payload programs,
we performed the same 28 indirect jumps and the final indirect jump with the
function pointer with two differences: first, the function pointer was cleared
from the cache (using the \texttt{clflush} instruction), and second, the
function pointer pointed to a different destination (specifically, to a
function
that measured the cache side channel, recording the indexes in the probe array
that appeared to be in the cache and then flushing all indexes).
At address 0x430000, we placed the code for our speculative gadget within the 
payload program, which is never directly called
(but executes speculatively). We note that this code must be loaded into the
cache but otherwise need not be accessed directly.

@@

We implemented a 2-symbol 5-state Busy Beaver Turing machine logic in a  
payload application. The gadget implementation reads a global state,
tape, and position, and determines how the state, tape, and position should be
updated. It encodes this update into a single byte, and loads from a global
buffer indexed by that byte, loading it into cache.
We randomized the stride length between indexes in
this array to avoid the CPU's stride prediction from preemptively loading values
in the buffer.

Our implementation is able to examine the state and current symbol on the Turing
tape and make decisions (in the speculative world) based on these values as long
as they are in cache. If one of these becomes uncached, one of two things will
happen. First, the speculative gadget may simply fail to report any result back
to the real world, as the uncached read takes too long to return before the
speculative world is squashed by the resolution of the function pointer. Second,
the CPU may \emph{speculate} on the value of the uncached item, which may be
incorrect. Of course, this does not affect the correctness of normal programs,
but it does impact the correctness of our speculative code. In particular, the
speculative gadget will compute using an incorrect value of state or current
symbol, and ultimately report the wrong update.

@@

We implemented the 5 state Busy Beaver configuration which requires over 47
million steps to complete. In order to have confidence the speculative gadget
executed with the correct inputs and reported the computed result correctly, we
repeat the execution a number of times and look for the value that was detected
as present in the cache most often over all the iterations. We measured the
error rate of our implementation as a function of how many redundant iterations
of the same step. Our Turing implementation discovered that 10 repetitions is
sufficient for the real world to determine the correct value, as the ``hit
rate'' for any one iteration observing the correct value loaded in cache was
over 80\%. Additionally we found that the speculative world could correctly
compute a million instructions with no errors at a rate of 324 instructions per
second.

@@

\subsection{AES Decryption}
\label{subsec:impl-aes}
The speculative world is able to take advantage of the AES-NI instructions to
decrypt messages. However, the speculative upper-limit of 220 instructions is not
enough to allow us to compute the key expansion, even using the
\texttt{aeskeygenassist} instructions. To avoid this, we preload the expanded
key schedule into the program, rather than the key. We note that an analyst
could observe the structure of a normal key schedule,
but we can avoid this by simply selecting 11 random round keys.
While this alternative key schedule variant is not ``true'' AES, we can easily
make the same change in our respective encryption function to make it compatible.
We note that
this should not weaken the security of AES, as we can ensure the round keys are
not linearly related.
We discuss further methods for obfuscating
the key in Section~\ref{subsec:nested-spec}.

@@

We demonstrate the speed that information can be decrypted via the speculative
world, and we vary the channel width of the side channel from 1 to 12 bits to
measure its performance. At low channel width, reading from the cache side
channel requires timing reads from only 2 locations, while at 12-bits, the side
channel requires reading $2^{12}$ locations. On the other hand, there is a fixed
overhead per speculative iteration that favors increased channel width to maximize
bandwidth. As shown in Figure~\ref{fig:spec_bandwidth}, 8 bits is the optimal
side channel width, allowing us to decrypt over 5,000 bits per second (625
Bytes/sec).
% Iterations: 20

\subsubsection{Obfuscating keys with nested speculation}
\label{subsec:nested-spec}

An analyst could locate the speculative entry point for our
decryption gadget by
searching for AES-NI instructions in the payload program's dead code, ultimately
discovering the keys that it uses. However, we find it is possible to overcome
this by having the trigger program communicate the decryption key to the payload
program via the branch predictor.

To accomplish this, we use multiple speculative entry points, each that derives a
different decryption key before calling a common decryption routine. Since
the exact speculative entry point is determined by the trigger program, an
analyst cannot discover the decryption key directly from the payload program.

An analyst could still enumerate all potential entry points, testing each one
until they find one that correctly decrypts the ciphertext. In a 1~MB binary,
there are (at most) only 1~million entry points, providing just 20~bits of
security.

To increase security, we instead use nested speculation to \emph{chain} entry points
together. Rather than derive
the key from a single entry point, we have each potential entry point perform
another indirect jump that the CPU cannot immediately resolve, forcing it to
speculate while already executing speculatively. The predicted target of that jump
will also be determined by the trigger program. For example, if the trigger
program makes 30 training jumps, followed by 10 additional indirect jumps, and
the payload program performs the same 30 training jumps before a stall, the CPU
will predict the payload program will also perform the next 10 jumps. If each
jump has the potential to land in up to 1024 possible locations, this would
provide in total $1024^{10} = 2^{100}$ possible keys, or 100~bits of security.
%We describe our implementation of this nested speculative execution concept in
%Section~\ref{subsec:nested-spec}.

@@

We implemented this technique, making 256 speculative landing spots, that each
shifted 8 unique bits into the 128-bit register \texttt{\%xmm0}, and then
performed an indirect jump. We then had the trigger program perform 16
indirect jumps that corresponded with 16 randomly-chosen landing spots in the
payload program, training the branch predictor. When the payload program reaches the
first speculative jump, it follows the same pattern speculatively, 
eventually filling
\texttt{\%xmm0} with the corresponding $16*8$ bits. We then used
the \texttt{aesenc} instruction to expand these 128-bits to a full
key schedule, and performed decryption as described previously.

@@

%Constructing an emulator making use of the speculative primitive requires  
%a trade off in expresive capability versus speed. 
We have implemented our custom instruction set architecture---SPASM---as a model
using two pseudo-registers, and
6-bit instruction length which allows for a relatively direct programming model
in which structured values can be entered into memory locations before making
a systemcall.

>>>>

\section{Implementation and Evaluation}
%We now discuss the details of our implementations and the information learned
%about speculative execution gained by these experiments. We discuss how we gain
%confidence in information passed back, and how a low number of iterations is
%sufficient to gain a strong confidence in results.

In this section, we discuss implementation details for our payload and trigger
programs.

%In each of our payload programs,
%we performed the same 28 indirect jumps and the final indirect jump with the
%function pointer with two differences: first, the function pointer was cleared
%from the cache (using the \texttt{clflush} instruction), and second, the
%function pointer pointed to a different destination (specifically, to a
%function
%that measured the cache side channel, recording the indexes in the probe array
%that appeared to be in the cache and then flushing all indexes).
%At address 0x430000, we placed the code for our speculative gadget within the 
%payload program, which is never directly called
%(but executes speculatively). We note that this code must be loaded into the
%cache but otherwise need not be accessed directly.
%

% Could move to an errata appendix?
%We randomized the stride length between indexes in
%this array to avoid the CPU's stride prediction from preemptively loading values
%in the buffer.

@@

We designed our Turing machine implementation to work with our custom trigger
program, with 28 indirect jumps mimicked by the Turing payload program.
We implemented a 2-symbol 5-state Busy Beaver Turing machine logic at the
speculative entry point (in 42 x86-64 instructions), returning the state update,
symbol to write, and tape move direction in a single byte via a cache side
channel.

We observed in our implementation that it is important that all values used in
the speculative world---as well as the code itself---be cached. If these are
evicted, the speculative code may fail to run, or the CPU may \emph{speculate}
on the value of the uncached item, which may be incorrect. While this does not
impact the correctness of normal programs whose incorrect speculations will be
resolved, our speculative code reports results back to the real world before
this resolution. In our Turing example, we observed this as incorrect state
transitions.

@@

%TODO: Ian please fill in these numbers
Instead, we
repeat the execution several times and look for the modal value over all
iterations. We measured the
error rate of our implementation as a function of how many redundant iterations
of the same step, and found that 10 redundant iterations resulted in 1 error every
million Turing steps, with the error rate dropping exponentially as iterations
increase. We choose TODO(Ian) iterations as a conservative bound (error rate measured
to be 0), and computed 1 million Turing steps at a rate of TODO(Ian)
steps per second.

% Was 324 steps per second, but was that at 10 iterations? But we just said
% 10 wasn't perfect (1 error in a million steps), so it seems confusing to say
% we computed a million steps without error at 10 iterations.

@@

\subsection{AES Decryption}
\label{subsec:impl-aes}
The speculative world is able to take advantage of the AES-NI instructions to
decrypt messages. However, the speculative upper-limit of about 175 $\mu$~ops is not
enough to allow us to compute the key expansion, even using the
\texttt{aeskeygenassist} instructions. To avoid this, we can either preload the
expanded key schedule into the program (instead of the key), or use a cheaper
(non-standard) key expansion algorithm.
For the former, we note that an analyst
could observe the structure of a normal key schedule,
but we can avoid this by simply selecting 11 random round keys.
We note that
this should not weaken the security of AES, as we can ensure the round keys are
not linearly related.

%We discuss further methods for obfuscating
%the key in Section~\ref{subsec:nested-spec}.
%

@@

% TODO: this could move elsewhere?
We demonstrate the speed that information can be decrypted via the speculative
world, and we vary the channel width of the side channel from 1 to 12 bits to
measure its performance. At low channel width, reading from the cache side
channel requires timing reads from only 2 locations, while at 12-bits, the side
channel requires reading $2^{12}$ locations. On the other hand, there is a fixed
overhead per speculative iteration that favors increased channel width to maximize
bandwidth. As shown in Figure~\ref{fig:spec_bandwidth}, 8 bits is the optimal
side channel width, allowing us to decrypt over 5,000 bits per second (625
Bytes/sec).
% Iterations: 20

@@

We also implemented our nested speculation technique for obfuscating keys,
making 256 speculative landing spots that each
shift 8 unique bits into the 128-bit register \texttt{\%xmm0}, and then
performing an indirect jump. We then had a custom trigger program perform 16
indirect jumps (after the initial 28) that corresponded with 16 randomly-chosen landing spots in the
payload program, training the branch predictor. When the payload program reaches the
first speculative jump, it follows the same pattern speculatively, 
eventually filling
\texttt{\%xmm0} with the corresponding $16*8$ bits. We then used
the \texttt{aesenc} instruction to expand these 128-bits to a full
key schedule, and performed decryption as described previously. Thus, without
the trigger program, an analyst has no information about what key is used to
decrypt the ciphertext in the payload.

@@

% TODO: needs a pass
%Constructing an emulator making use of the speculative primitive requires  
%a trade off in expresive capability versus speed. 
We have implemented our custom instruction set architecture---SPASM---as a model
using two pseudo-registers, and
6-bit instruction length which allows for a relatively direct programming model
in which structured values can be entered into memory locations before making
a systemcall.

### 2018-08-03 paper/payloads.tex
pass on implementation/payload reording

>>>>

\subsubsection{Obfuscating keys with nested speculation}
\label{subsec:nested-spec}

%TODO: ewust needs a pass to tie into previous section
An analyst could locate the speculative entry point for our
decryption gadget by
searching for AES-NI instructions in the payload program's dead code, ultimately
discovering the keys that it uses. However, we find it is possible to overcome
this by having the trigger program communicate the decryption key to the payload
program via the branch predictor.

To accomplish this, we use multiple speculative entry points, each that derives a
different decryption key before calling a common decryption routine. Since
the exact speculative entry point is determined by the trigger program, an
analyst cannot discover the decryption key directly from the payload program.

An analyst could still enumerate all potential entry points, testing each one
until they find one that correctly decrypts the ciphertext. In a 1~MB binary,
there are (at most) only 1~million entry points, providing just 20~bits of
security.

To increase security, we instead use nested speculation to \emph{chain} entry points
together. Rather than derive
the key from a single entry point, we have each potential entry point perform
another indirect jump that the CPU cannot immediately resolve, forcing it to
speculate while already executing speculatively. The predicted target of that jump
will also be determined by the trigger program. For example, if the trigger
program makes 30 training jumps, followed by 10 additional indirect jumps, and
the payload program performs the same 30 training jumps before a stall, the CPU
will predict the payload program will also perform the next 10 jumps. If each
jump has the potential to land in up to 1024 possible locations, this would
provide in total $1024^{10} = 2^{100}$ possible keys, or 100~bits of security.
%We describe our implementation of this nested speculative execution concept in
%Section~\ref{subsec:nested-spec}.
